Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-05-26 props.tex
comments on section I

\documentclass[conference]{IEEEtran}

@@

@@

@@

    Execution is by a step function \(m \xrightarrow{o} m' : m \in \mathcal{M} \rightarrow
    o \in \mathit{Obs} \times m' \in \mathcal{M}\).

    A policy tracks a mix of static and dynamic enforcement. A policy consists of
    a set of policy states, \(p \in \mathcal{P}\), a policy initializer
    \(pinit : \mathcal{M} \rightharpoonup \mathcal{P}\), and a policy-parameterized step function:
    \[\begin{split}
      m, p \xrightarrow{o}_P m', p' : m \in \mathcal{M} \times p \in \mathcal{P} & \rightharpoonup \\
      o \in \mathit{Obs} \times m' \in \mathcal{M} \times p' \in \mathcal{P} &
    \end{split}\]

    Axiomatically, if \(m,p \xrightarrow{o}_P m',p'\), then \(m \xrightarrow{o} m'\).
    The initializer is a partial function from machine states to policy
    states, reflecting that a policy may apply static constraints on the initial state. The parameterized
    step relation is likewise partial to allow for fail-stop behavior.

@@

    A machine trace \(M\) is a sequence of machine states \(m_0 m_1 m_2 \dots\) such that each
    pair of adjacent states are related by \(\longrightarrow\) with some observation. An
    {\it mptrace} is similarly a sequence of machine state-policy state pairs,
    \((m_0,\mathit{pinit}(m_0))(m_1,p_1) \dots\) such that adjacent pairs of pairs are related
    by \(\longrightarrow_P\). Steps being deterministic, we can relate an initial state to its trace
    via \(m \hookrightarrow M : m \in \mathcal{M} \rightarrow M \in \mathit{trace}\ \mathcal{M}\):

@@

    Since the step function is total, if \(m \hookrightarrow M\), then \(M\) must be infinite.
    We will also wish to take a prefix of the trace from an initial state up to when some condition
    holds. Let \(f\) be a predicate on machine states. Then we define \(m \hookrightarrow M | f\)
    (read ``\(M\) is the prefix from \(m\) until \(f\)'') where \(f\) may bind states in \(M\):
    \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1 \hookrightarrow M | f }
           {m_0 \hookrightarrow m_0 M | f}\]
    \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} f\ m_1}
           {m_0 \hookrightarrow m_0 m_1}\]

    We similarly define \(\hookrightarrow_P\), writing \(m_0,\mathit{pinit}(m_0) \longrightarrow \bot\) when
    there is no \(m_1,p_1\) such that \(m_0,\mathit{pinit}(m_0) \xrightarrow{o} m_1,p_1\) for any \(o\).
    \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} m_1,p_1 \hookrightarrow_P \MP}
           {m_0,p_0 \hookrightarrow_P m_0,p_0 \MP}\]
    \[\frac{m_0,p_0 \longrightarrow \bot}
           {m_0,p_0 \hookrightarrow_P m_0,p_0}\]

    With a policy we may see \(m,p \hookrightarrow \MP\) where \(\MP\) is finite due to a policy
    failure. We will identify the set of infinite traces as \(\top\), so \(\MP \in \top\) will
    indicate an infinite trace and \(\MP \not \in \top\) a finite one.
    We define a similar notion of taking the prefix until a condition:

@@

    \[T_1 \cdot T_2 \triangleq
      \begin{cases}
        T_2 & \text{if } T_1 = t \\
        t T_1' \cdot T_2 & \text{if } T_1 = t T_1' \\
      \end{cases}\]

    An observation trace \(O\) is the trace of observations from a machine or {\it mptrace},
    as defined by \(\mathit{Obs}\):
    \[\frac{M = mM' \hspace{.5cm} m \xrightarrow{o} m'}
           {\mathit{Obs}(M) = o\mathit{Obs}(M')}\]
    \[\frac{M = m}
           {\mathit{Obs}(M) = \tau}\]
    \[\frac{\MP = (m,p)\MP' \hspace{.5cm} m,p \xrightarrow{o} m',p'}
           {\mathit{Obs}(\MP) = o \mathit{Obs}(\MP')}\]
    \[\frac{\MP = m,p}
           {\mathit{Obs}(M) = \tau}\]

@@

@@

@@

@@

      The final state of a call sequence is a {\it call state}. If \(M\) is a call sequence with
      we map it to its call state, number of arguments, and the PC and SP values of its eventual
      return by the function \(\mathit{call} : \mathit{trace}\ \mathcal{M} \rightharpoonup \mathcal{M}
      \times \mathbb{N} \times \mathcal{W} \times \mathcal{W}\). We can also track a call sequence in the
      policy state, defining \(\mathit{call}_P : \mathcal{M} \times \mathcal{P} \rightharpoonup \mathbb{N}
      \times \mathcal{W} \times \mathcal{W}\) such that:

>>>>

\documentclass[conference]{IEEEtran}
\input{macros}

@@

\apt{important tex note: adopt/adapt macros from macros.tex to make it easier to switch notations if desire later.}

@@

\apt{Use set notation consistently rather than mixing in grammar notation ($::=$ and $|$).}

@@

    Execution is by a step function \(m \xrightarrow{o} m' : m \in \mathcal{M} \rightarrow
    o \in \mathit{Obs} \times m' \in \mathcal{M}\).\apt{Very awkward: don't repeat notation conventions in type.}

    A policy tracks a mix of static and dynamic enforcement.\apt{??} A policy consists of
    a set of policy states, \(p \in \mathcal{P}\), a policy initializer
    \(pinit : \mathcal{M} \rightharpoonup \mathcal{P}\), and a policy-parameterized step function:
    \[\begin{split}
      m, p \xrightarrow{o}_P m', p' : m \in \mathcal{M} \times p \in \mathcal{P} & \rightharpoonup \\
      o \in \mathit{Obs} \times m' \in \mathcal{M} \times p' \in \mathcal{P} &
    \end{split}\]
    \apt{I would suggest parenthesizing the pairs $(m,p)$ unless that looks too busy. (This sort of experiment is exactly why macros are a good idea.)}

    Axiomatically, if \(m,p \xrightarrow{o}_P m',p'\), then \(m \xrightarrow{o} m'\).
    \apt{Why not follow coq development by defining policy step function $\mathcal{M} \times \mathcal{P} \rightarrow \mathcal{P}$ separately and then \emph{defining}
      combined function $\mathcal{M} \times \mathcal{P} \rightarrow \mathcal{M} \times \mathcal{P}$.}
    The initializer is a partial function from machine states to policy
    states, reflecting that a policy may apply static constraints on the initial state. The parameterized
    step relation is likewise partial to allow for fail-stop behavior.

@@

    A machine trace \(M\) is a sequence of machine states \(m_0 m_1 m_2 \dots\) such that each
    pair of adjacent states are related by \(\longrightarrow\) with some observation. An
    {\it mptrace} is similarly a sequence of machine state-policy state pairs,
    \((m_0,\mathit{pinit}(m_0))(m_1,p_1) \dots\) \apt{why does it need to start with $\mathit{pinit}$?}
    such that adjacent pairs of pairs are related
    by \(\longrightarrow_P\). Steps being deterministic\apt{awkward}, we can relate an initial state to its trace
    via \(m \hookrightarrow M : m \in \mathcal{M} \rightarrow M \in \mathit{trace}\ \mathcal{M}\):

@@

    \apt{Note that this is a coinductive relation.}

    Since the step function is total, if \(m \hookrightarrow M\), then \(M\) must be infinite.
    We will also wish to take a prefix of the trace from an initial state up to when some condition
    holds\apt{``up through the state where some first condition first holds'' (?)}.
    Let \(f\) be a predicate on machine states. Then we define \(m \hookrightarrow M | f\)
    (read ``\(M\) is the prefix from \(m\) until \(f\)'') where \(f\) may bind\apt{wrong word, and this phrase is redundant anyhow} states in \(M\):
    \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} \neg f\ m_1 \hspace{.5cm} m_1 \hookrightarrow M | f }
           {m_0 \hookrightarrow m_0 M | f}\]
    \[\frac{m_0 \xrightarrow{o} m_1 \hspace{.5cm} f\ m_1}
           {m_0 \hookrightarrow m_0 m_1}\]

           \apt{What if $f$ holds on $m_0$ ? }

    We similarly define \(\hookrightarrow_P\), writing \(m_0,\mathit{pinit}(m_0) \longrightarrow \bot\) when
    there is no \(m_1,p_1\) such that \(m_0,\mathit{pinit}(m_0) \xrightarrow{o} m_1,p_1\) for any \(o\).
    \apt{What does $\mathit{pinit}$ have to do with it?. In any case, suggest using $\nrightarrow$ or similar.}
    \[\frac{m_0,p_0 \xrightarrow{o} m_1,p_1 \hspace{.5cm} m_1,p_1 \hookrightarrow_P \MP}
           {m_0,p_0 \hookrightarrow_P m_0,p_0 \MP}\]
    \[\frac{m_0,p_0 \longrightarrow \bot}
           {m_0,p_0 \hookrightarrow_P m_0,p_0}\]

    With a policy we may see \(m,p \hookrightarrow \MP\) where \(\MP\) is finite due to a policy
    failure. We will identify the set of infinite traces \apt{over what? Don't think set is the right notion here.}
    as \(\top\), so \(\MP \in \top\) will
    indicate an infinite trace and \(\MP \not \in \top\) a finite one.
    We define a similar notion of taking the prefix until a condition:

@@

    \apt{If we're defining $\mathit{trace}$ generically, why not do so earlier to avoid multiple definitions of $\hookrightarrow$ etc.?}
    \[T_1 \cdot T_2 \triangleq
      \begin{cases}
        t T_2 & \text{if } T_1 = t \\
        t T_1' \cdot T_2 & \text{if } T_1 = t T_1' \\
      \end{cases}\]

      \apt{But actually, the coq definition is defined for all $T_1$ (even if that isn't very useful). And isn't it important to let $T_2$ be empty?}
    An observation trace \(O\) is the trace of observations from a machine or {\it mptrace},
    as defined by \(\mathit{Obs}\):
    \[\frac{M = mM' \hspace{.5cm} m \xrightarrow{o} m'}
           {\mathit{Obs}(M) = o\mathit{Obs}(M')}\]
    \[\frac{M = m}
           {\mathit{Obs}(M) = \tau}\]
    \[\frac{\MP = (m,p)\MP' \hspace{.5cm} m,p \xrightarrow{o} m',p'}
           {\mathit{Obs}(\MP) = o \mathit{Obs}(\MP')}\]
    \[\frac{\MP = m,p}
           {\mathit{Obs}(M) = \tau}\]

@@

    \apt{This is not quite the same as the coq definition. Is it really equivalent?}

@@

    \apt{Ditto.}

@@

           \apt{Does this notion appear in Coq? It seems potentially useful, but it seems to require
             knowing whether we are in the finite or infinite case.}

@@

      \apt{But the \emph{properties} don't need these notions!}

      The final state of a call sequence \apt{what's that ??} is a {\it call state}. If \(M\) is a call sequence with
      we map it \apt{??} to its call state, number of arguments, and the PC and SP values of its eventual
      return by the function \(\mathit{call} : \mathit{trace}\ \mathcal{M} \rightharpoonup \mathcal{M}
      \times \mathbb{N} \times \mathcal{W} \times \mathcal{W}\). We can also track a call sequence in the
      policy state, defining \(\mathit{call}_P : \mathcal{M} \times \mathcal{P} \rightharpoonup \mathbb{N}
      \times \mathcal{W} \times \mathcal{W}\) such that:
